\Large \bf 6.824 Final Project

### 2014-05-12 writeup/paper.tex
added to writeup

\textbf{Single server crashing and restarting, disk in-tact}: The
recovery process contacts a random replica in the same replica group.
The recovering server tells the up-to-date server its current state,
and the up-to-date server responds with any state the recoverer is
missing. This works because the server only discards old decisions if
all participants have marked that sequence number as 'done'.

\textbf{Single server crashing and disk lost}: The recovery process
contacts a random replica in the same replica group, and tells the
up-to-date server that it has no state. The two servers open an
auxillary TCP connection to send over the relevant portions of the
database. This avoids RPC overhead, since the database might be large.
The database has an in-order iterator, in case the TCP connection
fails, the recoverer can re-establish the connection and ask the
sender to pick up where it left off. If the sending server fails, then
the recovering server asks a new replica. This allows for a failure
mode where the original server remians locked permanently. We have
decided that this is preferable to sacrificing consistency, and the
system administartor can resolve the problem by rebooting the hing
server.

\textbf{Complete crash of all servers}: If all servers lose their disk
state, then all will restart and begin like a brand new deployment. If
the disks are not lost, then the recovery process will read the paxos
state written to disk and pick up where it left off. Because we write
to disk before any operation commits, data is not lost.

%% \subsubsection{Memory and Performance Considerations}
%% There are a number of design choices which were made in order to address both memory management and overall performance.

%% Firstly, the service must be able to handle database content which is too large to fit in volatile memory.  Towards this end, values are written to disk immediately as described above so it does not need to be kept in memory.  More importantly, however, servers that need to transfer shards (either because the shard master decided a shard needs to move or because a server is recovering) must be able to do so in multiple RPCs since memory is too small to create a reply with the entire database.  Therefore, a shard transfer occurrs with the following protocol:
%% \begin{itemize}
%% \item The requester sends a request to a peer, including a request number, and the current configuration number.  If doing a shard transfer, the shard number is also included.
%% \item If the receiver is currently at the appropriate configuration, it accepts the request.  It will then begin populating a reply with client response data, a list of seen operations, and the desired key/value data.  While doing this, it will periodically check memory usage; if it exceeds a configurable threshold, it will reply with whatever it currently has and mark the reply as incomplete.
%% \item The requester copies the data from the reply to its disk.  If the reply was marked as incomplete, it sends another request to the same peer with an incremented request number.
%% \item This continues until the reply is marked as complete, at which point the requester launches a goroutine to make sure the sender receives an acknowledgement of the receipt of the final data message.
%% \item If at any point the requester decides that the contacted peer is dead, it launches an acknowledgement goroutine as previously described for that peer (to terminate the transfer in case the peer wakes up or is not dead) and moves on to try a different peer, resetting the request number to 0.
%% \end{itemize}
%% Both the requester and the sender ignore client requests until the transfer is complete - this ensures that the requester will receive a consistent snapshot of the data (and is the reason the requester launches acknowledgement goroutines since otherwise a peer erroneously declared dead would be halted forever).

%% One aspect which addresses performance is that progress can still be made while recovery is taking place as long as a majority of servers are not recovering.  For example, if a replica group has three servers and a shard transfer is ocurring between two of them (perhaps because one just restarted), a client can still have its requests processed if it contacts the third peer.  This is because the Paxos service runs independently of the shard key/value service; although the latter will ignore client requests on the transferring servers, their Paxos processes can still respond and make progress.  Since the Paxos instances are stored persistently on disk, the two occupied servers can catch themselves up from the log when they complete the transfer.

%% In addition, effort was made to utilize memory in addition to the disk in order to increase performance.  The LevelDB database has the option to use a least-recently-used (LRU) cache to store recently used entries in memory.  In addition, a configuration variable was added to the service to allow values to be written to memory as well as disk until the memory is full.  These two options, combined with the native cache of the operating system, allow many needed values to be read quickly from memory rather than always consulting the disk.

%% When one server contacts another via the FetchRecovery RPC, both
%% servers enter a recovering state. This state means that neither are
%% allowed to modify their databases \emph{except} through the recovery
%% process. Depending on how many replicas there are, this may render a
%% replica group unable to process requests while recovery is underway.
%% If there are enough replicas so that paxos majority can still be
%% reached, then the recovering servers will catch up once they come back
%% online, using the functionality already built in to sync lagging paxos
%% servers.

%% All recovery processes send over both the relevant portions of the k/v
%% store, as well as paxos state and any state needed to detect
%% duplicates.

%% We use leveldb. To guarantee persistence, we enforce writing to disk
%% all operations before returning. This slows performance down, but
%% prevents loss of data that the client wants committed. Speed could be
%% gained using a Harp-like set of pointers, but this would be
%% complicated and the client would have to understand the modified
%% persistence model.

@@

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{
\includegraphics[scale=.4]{recovery.png}}
\caption{Performance increases dramatically as the number of keys is reduced (and the size of entries is increased).}
\label{fig:recovery}
\end{figure}

@@

One notable bottleneck is disk reading and writing.  The database library wrapper, LeveDB, offers a cache as well as compression and is quite fast.  However, as discussed previously, benchmark tests revealed that scanning the database to compile data for shard transfers accounted for a significant portion of total transfer time.  Additionally, in the process of debugging networked RPCs, simple logging was implemented which slowed down operations by nearly a factor of ten; this was a good demonstration of disk latency and how care must be taken in even mundane aspects of the system.  In the future, the system topology could be somewhat altered to allow for more parallelizable recovery and transfer procedures to help minimize the effect of disk latency (for example, a topology inspired by FDS).

>>>>

\textbf{Single server crashing and restarting, disk in-tact or lost}:
The server will contact its peers to figure out the state it is
missing, and it will then send RPCs to collect the missing state. The
server it is doing state transfer with cannot change disk state until
the entire process is done. If the disk is lost, the request indicates
that it has no state, which causes the entire database to be sent.

\textbf{Complete crash of all servers}: Servers read their state from
disk, and then contact peers to figure out if they are out of date.

%% \subsubsection{Memory and Performance Considerations}
%% There are a number of design choices which were made in order to
%% address both memory management and overall performance.

%% Firstly, the service must be able to handle database content which
%% is too large to fit in volatile memory.  Towards this end, values
%% are written to disk immediately as described above so it does not
%% need to be kept in memory.  More importantly, however, servers that
%% need to transfer shards (either because the shard master decided a
%% shard needs to move or because a server is recovering) must be able
%% to do so in multiple RPCs since memory is too small to create a
%% reply with the entire database.  Therefore, a shard transfer
%% occurrs with the following protocol:
%% \begin{itemize}
%% \item The requester sends a request to a peer, including a request
%%   number, and the current configuration number.  If doing a shard
%%   transfer, the shard number is also included.
%% \item If the receiver is currently at the appropriate
%%   configuration, it accepts the request.  It will then begin
%%   populating a reply with client response data, a list of seen
%%   operations, and the desired key/value data.  While doing this, it
%%   will periodically check memory usage; if it exceeds a
%%   configurable threshold, it will reply with whatever it currently
%%   has and mark the reply as incomplete.
%% \item The requester copies the data from the reply to its disk.  If
%%   the reply was marked as incomplete, it sends another request to
%%   the same peer with an incremented request number.
%% \item This continues until the reply is marked as complete, at
%%   which point the requester launches a goroutine to make sure the
%%   sender receives an acknowledgement of the receipt of the final
%%   data message.
%% \item If at any point the requester decides that the contacted peer
%%   is dead, it launches an acknowledgement goroutine as previously
%%   described for that peer (to terminate the transfer in case the
%%   peer wakes up or is not dead) and moves on to try a different
%%   peer, resetting the request number to 0.
%% \end{itemize}
%% Both the requester and the sender ignore client requests until the
%% transfer is complete - this ensures that the requester will receive
%% a consistent snapshot of the data (and is the reason the requester
%% launches acknowledgement goroutines since otherwise a peer
%% erroneously declared dead would be halted forever).

One aspect which addresses performance is that progress can still
be made while recovery is taking place as long as a majority of
servers are not recovering.  For example, if a replica group has
three servers and a shard transfer is ocurring between two of them
(perhaps because one just restarted), a client can still have its
requests processed if it contacts the third peer.  This is because
the Paxos service runs independently of the shard key/value
service; although the latter will ignore client requests on the
transferring servers, their Paxos processes can still respond and
make progress.  Since the Paxos instances are stored persistently
on disk, the two occupied servers can catch themselves up from the
log when they complete the transfer.

In addition, effort was made to utilize memory in addition to the
disk in order to increase performance.  The LevelDB database has
the option to use a least-recently-used (LRU) cache to store
recently used entries in memory.  In addition, a configuration
variable was added to the service to allow values to be written to
memory as well as disk until the memory is full.  These two
options, combined with the native cache of the operating system,
allow many needed values to be read quickly from memory rather than
always consulting the disk.

When one server contacts another via the FetchRecovery RPC, both
servers enter a recovering state. This state means that neither are
allowed to modify their databases \emph{except} through the
recovery process. Depending on how many replicas there are, this
may render a replica group unable to process requests while
recovery is underway.  If there are enough replicas so that paxos
majority can still be reached, then the recovering servers will
catch up once they come back online, using the functionality
already built in to sync lagging paxos servers.

All recovery processes send over both the relevant portions of the
k/v store, as well as paxos state and any state needed to detect
duplicates.

%% We use leveldb. To guarantee persistence, we enforce writing to
%% disk all operations before returning. This slows performance down,
%% but prevents loss of data that the client wants committed. Speed
%% could be gained using a Harp-like set of pointers, but this would
%% be complicated and the client would have to understand the modified
%% persistence model.

@@

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{
\includegraphics[scale=.7]{recovery.png}}
\caption{Performance increases dramatically as the number of keys is reduced (and the size of entries is increased).}
\label{fig:recovery}
\end{figure}

@@

One notable bottleneck is disk reading and writing.  The database
library wrapper, LeveDB, offers a cache as well as compression and is
quite fast.  However, as discussed previously, benchmark tests
revealed that scanning the database to compile data for shard
transfers accounted for a significant portion of total transfer time.
Additionally, in the process of debugging networked RPCs, simple
logging was implemented which slowed down operations by nearly a
factor of ten (it was disabled for benchmarks); this was a good
demonstration of disk latency and how care must be taken in even
mundane aspects of the system.  In the future, the system topology
could be somewhat altered to allow for more parallelizable recovery
and transfer procedures to help minimize the effect of disk latency
(for example, a topology inspired by FDS).

For disk recovery, we also examined the possibility of using a raw TCP
stream to send the data, instead of using RPCs, in order to reduce
overhead. The transfer rate for a 1GB database was about 4.5
Mbps. This was approximately equivalent to the data rate at which the
database itself was created locally. Thus, the overhead necessary to
read and write values to the database is a key bottleneck.
